Lattice Boltzmann study of mass transfer for complicated geometries

### 2012-02-06 Text/introduction.tex
Started doing showcases for the second part of the paper.

\section{Mass transfer definitions}
By the definition the mass transfer coefficient from the surface with the imposed constant
concentration $\cbubble$ is the following:
\beq
k_L=\frac{\dot{m}}{P \Delta C},
\feq
where $\dot{m}$ is the mass rate $\Bigl[\frac{kg}{s}\Bigr]$, $P$ is the area of the surface
$\Bigl[m^2\Bigr]$, $\Delta C$ is the concentration difference between the surface and the surrounding medium
$\Bigl[\frac{kg}{m^3}\Bigr]$. Therefore, $k_L$ has a dimension of velocity
$\Bigl[\frac{m}{s}\Bigr]$. Usually, the surrounding medium concentration is taken at the infinite distance
from the bubble. However, in the case of complicated geometries and non-homogeneous concentrations, 
the medium concentration can be the average concentration in the domain, the flux averaged
concentration at the inlet or outlet. Thus, one needs to establish a thorough definition of the volumetric
mass transfer coefficient in the case of complex geometries and non-trivial hydrodynamic velocity patterns.

@@

\subsection{$Ca=0.097$ results}
To understand the process happening in unit cells one need to keep a track of a few
characteristics. We chose following representative characteristics: the average concentration in
the unit cell with time, the concentration difference between adjacent unit cells. The latter is
the interesting characteristics which shows how well slug liquid is mixed (to be shown later). The
argument behind this characteristic is that if the following situation is fulfilled: the continuity
equation is satisfied then the concentration along the coordinate changes according to Eq.
\ref{main:mass:transfer:expression}:
\beq
C(x)= \cstar \Bigl(1-\exp\bigl(-\vol \frac{x}{\uliq+\ugas} \bigr)\Bigr)
\feq
%After the transition to the reference frame moving with the velocity $\ububble$, one will have the
%same continuous picture but the concentration dependency will change since velocity is changed as
%well:
%\beq
%C(x)= \cstar \Bigl(1-\exp\bigl(-\vol \frac{x}{\uliq+\ugas-\ububble} \bigr)\Bigr),
%\feq
%Thus, the volumetric mass transfer coefficient for two concentrations which are located at the
%distance $\lunit$ can be found as:
%\beqal
%&\vol \frac{\lunit}{\uliq+\ugas-\ububble}=\ln\Bigl(\frac{\cstar-C_1}{\cstar-C_2}\Bigr)\\
%&\vol
%\frac{\lunit}{\uliq+\ugas}=\frac{\uliq+\ugas-\ububble}{\uliq+\ugas}\ln\Bigl(\frac{\cstar-C_1}{
%\cstar-C_2}\Bigr)
%\feqal
%Parameters $\frac{\uliq+\ugas-\ububble}{\uliq+\ugas}$ were calculated and for representative
%Capillary numbers equal to $0.11,0.04,-0.02,-0.06,-0.10$.
Therefore, the following correlation was examined between adjacent cells:
\beqal
\label{volumetric:between:adjacent:cells}
&C_1=C(x)=C^* \Bigl(1-e^{-\vol \frac{x}{\ugas+\uliq}}\Bigr)\\
&C_2=C(x+L)=C^* \Bigl(1-e^{-\vol \frac{x+L}{\ugas+\uliq}}\Bigr)\\
&k_L\frac{L}{\ugas+\uliq}=\frac{C^*-C_1}{C^{*}-C_2}
\feqal
Fig. \ref{fig:unit:4} shows one of the examples for velocity scale $20$ and  $4$ unit numbers (all
scales produce the same results) for the volumetric mass transfer coefficient based on difference
between two adjacent cells concentration averages. One can see that this number is different in
comparison with the number used for periodic boundary
conditions. Moreover, it is changed from segment to segment. The same dependencies can be found for
$6$ and $8$ unit cells simulations but we do not present them here.

However, if one wants to examine how the average concentration changes in time according to Eq.
\ref{theor:continuous:mass:transfer} for different unit cells then one will see interesting
thing for the volumetric mass transfer correlation, see Fig. \ref{fig:moving:average:ca0097}:
\beq
\label{moving:average}
\vol\frac{\lunit}{\ugas+\uliq}=\frac{\lunit}{\ububble (t_2
-t_1)}\ln\Bigl(\frac{C^*-C_1}{C^*-C_2}\Bigr)
\feq
\begin{figure}[htb!]
%\includegraphics[width=0.5\textwidth]{Figures/aver_units4scale10.eps}\hfill
%\includegraphics[width=0.5\textwidth]{Figures/coeff_units4scale10.eps}\\
%\includegraphics[width=0.5\textwidth]{Figures/aver_units4scale15.eps}\hfill
%\includegraphics[width=0.5\textwidth]{Figures/coeff_units4scale15.eps}\\
\includegraphics[width=0.5\textwidth]{Figures/aver_units4scale20.eps}\hfill
\includegraphics[width=0.5\textwidth]{Figures/coeff_units4scale20.eps}\\
\caption{Average concentrations and volumetric coefficients for $4$ units cells. The volumetric
mass transfer coefficient is calculated based on the difference between averages for different unit
cells.
\label{fig:unit:4}}
\end{figure}
\begin{figure}[htb!]
\includegraphics[width=0.5\textwidth]{Figures/aver_moving_window4scale10.eps}
\includegraphics[width=0.5\textwidth]{Figures/aver_moving_window6scale10.eps}\\
\includegraphics[width=0.5\textwidth]{Figures/aver_moving_window8scale10.eps}
\caption{The non-dimensional volumetric mass transfer coefficient defined in Eq.
\ref{moving:average} for $4$ (top left),$6$ (top right),$8$ cell units (bottom). Only scale
$10$ is presented since all other simulations produce the same results. One can see that $4$ unit
cells is not enough to see avoid the influence of boundaries. However, the results for $6$ and $8$
unit cells are consistent and show that beginning from third unit cell the results and ending with
prelast cell results are consistent with periodic boundary simulations and
\citet{vanbaten-circular} formulations.
\label{fig:moving:average:ca0097}}
\end{figure}

\subsection{$Ca=1.040$ results}
The same correlations were examined for the different velocity pattern $Ca=1.040$ and initial
velocity $\ububble=0.055$ and initial diffusion $D=0.0008375$. A few unit cells simulation for this
capillary number are unstable. To improve stability we changed original Peclet number by increasing
diffusion to reveal all the correlations. The results for unit cells are the same as for $Ca=0.097$:
at least $6$ unit cells are required to be simulated to avoid the influence of boundaries. Thus,
only $6$ unit cells results are presented in Fig. \ref{fig:6:units:ca1040}.One can see that the
volumetric mass transfer coefficient based
on the difference between unit cells averages is constant for internal
unit cells. This case is different from the previous case where those coefficients were not equal
to each other. The average concentration being constant means that there is no backward convection
flux and there is no slug tracer mixing. This test is a good example to indicate whether there is
mixing in the liquid slug or not.
\begin{figure}[htb!]
\includegraphics[width=0.5\textwidth]{Figures/aver_units6scaleu2scaled5.eps}
\includegraphics[width=0.5\textwidth]{Figures/novortex6scaleu2scaled5.eps}\\
\caption{Results for 6 unit cells. The Peclet number equals to $Pe=2644$.
One can see that average concentrations reach certain number and stay.
That means whatever was generated by medium is transfered through
boundary. One can see that the volumetric mass transfer coefficient based
on the difference between unit cells averages is constant for internal
unit cells. This is the indication that the tracer in the slug is not
mixed.\label{fig:6:units:ca1040}}
\end{figure}
The volumetric mass transfer coefficient with periodic boundary simulations for $Pe=2644$ (velocity
scale $2$, diffusion scale $10$) and time dependent, Eq. \ref{theor:continuous:mass:transfer},
average volumetric mass transfer coefficient are presented in Fig. \ref{fig:periodic:ca1040}. Note
that they coincide but are different from calculated coefficients based on spatial location.
Therefore, the time based simulations do not readily convert to spatial based simulations.
\begin{figure}
\includegraphics[width=0.5\textwidth]{Figures/volume_ca_104_scaleu2scaled5.eps}
\includegraphics[width=0.5\textwidth]{Figures/aver_moving_window6scaleu2scaled5.eps}\\
\caption{The periodic (left) and time unit cells averaging volumetric mass transfer coefficients.
One can see that they coincide, but are different from spatial based volumetric mass transfer
coefficient, see Fig. \ref{fig:6:units:ca1040} \label{fig:periodic:ca1040}}
\end{figure}

@@

domain concentration through any formulation as they produce consistent results. Formulation of

@@

have indicator of how well the liquid slug is mixed. Finally, a sample of results was

>>>>

\section{Mass transfer definitions}
By the definition the mass transfer coefficient from the surface with the imposed constant
concentration $\cbubble$ is the following:
\beq
\label{eq:main:definition}
k_L=\frac{\dot{m}}{P \Delta C},
\feq
where $\dot{m}$ is the mass rate $\Bigl[\frac{kg}{s}\Bigr]$, $P$ is the area of the surface
$\Bigl[m^2\Bigr]$, $\Delta C$ is the concentration difference between the surface and the surrounding medium
$\Bigl[\frac{kg}{m^3}\Bigr]$. Therefore, $k_L$ has a dimension of velocity
$\Bigl[\frac{m}{s}\Bigr]$. Usually, the surrounding medium concentration is taken at the infinite distance
from the bubble. However, in the case of complicated geometries and non-homogeneous concentrations, 
the medium concentration can be the average concentration in the domain, the flux averaged
concentration at the inlet or outlet. Thus, one needs to establish a thorough definition of the volumetric
mass transfer coefficient in the case of complex geometries and non-trivial hydrodynamic velocity patterns.

@@

\subsection{$Ca=0.097$ results}
To understand the process happening in unit cells one need to keep a track of a few
characteristics. We chose following representative characteristics: the average concentration in
the unit cell with time, the concentration difference between adjacent unit cells. The latter is
the interesting characteristics which shows how well slug liquid is mixed (to be shown later). The
argument behind this characteristic is that if the following situation is fulfilled: the continuity
equation is satisfied then the concentration along the coordinate changes according to Eq.
\ref{main:mass:transfer:expression}:
\beq
C(x)= \cstar \Bigl(1-\exp\bigl(-\vol \frac{x}{\uliq+\ugas} \bigr)\Bigr)
\feq
%After the transition to the reference frame moving with the velocity $\ububble$, one will have the
%same continuous picture but the concentration dependency will change since velocity is changed as
%well:
%\beq
%C(x)= \cstar \Bigl(1-\exp\bigl(-\vol \frac{x}{\uliq+\ugas-\ububble} \bigr)\Bigr),
%\feq
%Thus, the volumetric mass transfer coefficient for two concentrations which are located at the
%distance $\lunit$ can be found as:
%\beqal
%&\vol \frac{\lunit}{\uliq+\ugas-\ububble}=\ln\Bigl(\frac{\cstar-C_1}{\cstar-C_2}\Bigr)\\
%&\vol
%\frac{\lunit}{\uliq+\ugas}=\frac{\uliq+\ugas-\ububble}{\uliq+\ugas}\ln\Bigl(\frac{\cstar-C_1}{
%\cstar-C_2}\Bigr)
%\feqal
%Parameters $\frac{\uliq+\ugas-\ububble}{\uliq+\ugas}$ were calculated and for representative
%Capillary numbers equal to $0.11,0.04,-0.02,-0.06,-0.10$.
In the case of present simulations the situation is different as the reference frame is moving with
the bubble. Thus, the correlation between adjacent cells as follows:
\beqal
\label{volumetric:between:adjacent:cells}
&C_1=C(x)=C^* \Bigl(1-e^{-\vol \frac{x}{\ugas+\uliq-\ububble}}\Bigr)\\
&C_2=C(x+L)=C^* \Bigl(1-e^{-\vol \frac{x+L}{\ugas+\uliq-\ububble}}\Bigr)\\
&k_L\frac{L}{\ugas+\uliq-\ububble}=\frac{C^*-C_1}{C^{*}-C_2}\\
&k_L\frac{L}{U}=\frac{C^*-C_1}{C^*-C_2},
\feqal
where $U=\ugas+\uliq-\ububble$. We will see later that this characteristic is really important to
see whether the liquid slug is mixed. Fig. \ref{fig:unit:4} shows one of the examples for velocity
scale $20$ and  $4$ unit numbers (all
scales produce the same results) for the volumetric mass transfer coefficient $k_L \frac{L}{U}$
(different from $\volnondim$) based on difference
between two adjacent cells concentration averages. One can see that this number is changed from
segment to segment. The same dependencies can be found for
$6$ and $8$ unit cells simulations but we do not present them here.

However, if one wants to examine how the average concentration changes in time according to Eq.
\ref{theor:continuous:mass:transfer} for different unit cells then one will see interesting
thing for the volumetric mass transfer correlation, see Fig. \ref{fig:moving:average:ca0097}:
\beq
\label{moving:average}
\vol\frac{\lunit}{\ugas+\uliq}=\frac{\lunit}{\ububble (t_2
-t_1)}\ln\Bigl(\frac{C^*-C_1}{C^*-C_2}\Bigr)
\feq
\begin{figure}[htb!]
%\includegraphics[width=0.5\textwidth]{Figures/aver_units4scale10.eps}\hfill
%\includegraphics[width=0.5\textwidth]{Figures/coeff_units4scale10.eps}\\
%\includegraphics[width=0.5\textwidth]{Figures/aver_units4scale15.eps}\hfill
%\includegraphics[width=0.5\textwidth]{Figures/coeff_units4scale15.eps}\\
\includegraphics[width=0.5\textwidth]{Figures/aver_units4scale20.eps}\hfill
\includegraphics[width=0.5\textwidth]{Figures/coeff_units4scale20.eps}\\
\caption{Average concentrations (left) and volumetric coefficients (right) $\vol \frac{L}{U}$ (
different from $\volnondim$) for $4$ units cells. The volumetric
mass transfer coefficient is calculated based on the difference between averages for different unit
cells. One can see that the numbers for the volumetric mass transfer coefficient are different for
different unit cells. This is the indication that the liquid slug is mixed and the spatial location
does not easily transfers to time domain. 
\label{fig:unit:4}}
\end{figure}
\begin{figure}[htb!]
\includegraphics[width=0.5\textwidth]{Figures/aver_moving_window4scale10.eps}
\includegraphics[width=0.5\textwidth]{Figures/aver_moving_window6scale10.eps}\\
\includegraphics[width=0.5\textwidth]{Figures/aver_moving_window8scale10.eps}
\caption{The non-dimensional volumetric mass transfer coefficient defined in Eq.
\ref{moving:average} for $4$ (top left),$6$ (top right),$8$ cell units (bottom). Only scale
$10$ is presented since all other simulations produce the same results. One can see that $4$ unit
cells is not enough to avoid the influence of boundaries. However, the results for $6$ and $8$
unit cells are consistent and show that beginning from third unit cell the results and ending with
prelast cell results are consistent with periodic boundary simulations and
\citet{vanbaten-circular} formulations.
\label{fig:moving:average:ca0097}}
\end{figure}

\subsection{$Ca=1.040$ results}
The same correlations were examined for the different velocity pattern $Ca=1.040$ and initial
velocity $\ububble=0.055$ and initial diffusion $D=0.0008375$. A few unit cells simulation for
this
capillary number are unstable. To improve stability we changed original Peclet number by increasing
diffusion to reveal all the correlations. The results for unit cells are the same as for $Ca=0.097$:
at least $6$ unit cells are required to be simulated to avoid the influence of boundaries. Thus,
only $6$ unit cells results are presented in Fig. \ref{fig:6:units:ca1040}. As well the volumetric
mass transfer coefficient, Eq. \ref{volumetric:between:adjacent:cells}, is shown in Fig.
\ref{fig:6:units:ca1040}. One can see that the average volume concentration, as well the volumetric
mass transfer coefficients, come to the constant and are not increasing with time. Thus, all the
mass generated with medium is transfered through the boundaries. This is the indication that the
liquid slug is not mixed. Thus, the volumetric mass transfer coefficient $\volnondim$ can be
calculated according to the definition, Eq. \ref{eq:main:definition}:
\beq
\label{inlet:outlet:spatial:location}
\vol = \frac{\dot{m}}{V \Delta C},
\feq 
where $V$ is the unit cell volume, $\Delta C$ is the concentration difference between the bubble
surface concentration $\cstar$ and the average medium concentration $C_{aver}$. Note that in this
case one can calculate the mass generation as:
\beq
\dot{m}=(\coutlet-\cinlet)\uoutlet,
\feq
where $\coutlet$ and $\cinlet$ are taken for the current unit. Fig. \ref{fig:periodic:ca1040}
(bottom) shows the volumetric mass transfer coefficient, based on spatial calculations of
inlet/outlet concentrations. One can see that the volumetric mass transfer coefficient is close to
the calculated volumetric mass transfer coefficient using the time averaged approach and periodic
boundaries one unit cells simulations. 
\begin{figure}[htb!]
\includegraphics[width=0.5\textwidth]{Figures/aver_units6scaleu2scaled5.eps}
\includegraphics[width=0.5\textwidth]{Figures/novortex6scaleu2scaled5.eps}\\
\caption{Results for 6 unit cells. The Peclet number equals to $Pe=2644$.
One can see that average concentrations reach certain number and stay.
That means whatever was
generated by medium is transfered through the outlet
boundary. One can see that the volumetric mass transfer coefficient based
on the difference between unit cells averages, Eq. \ref{volumetric:between:adjacent:cells}, is
constant for internal
unit cells. This is the indication that the tracer in the slug is not
mixed and the volumetric mass transfer coefficient, $\volnondim$, can be
calculated using the spatial approach, see Fig.
\ref{fig:periodic:ca1040}.\label{fig:6:units:ca1040}}
\end{figure}
The volumetric mass transfer coefficient with periodic boundary simulations for $Pe=2644$ (velocity
scale $2$, diffusion scale $10$) and time dependent, Eq. \ref{theor:continuous:mass:transfer},
average volumetric mass transfer coefficient are presented in Fig. \ref{fig:periodic:ca1040}. Note
that they coincide but are different from calculated coefficients based on spatial location.
Therefore, the time based simulations do not readily convert to spatial based simulations.
\begin{figure}
\includegraphics[width=0.5\textwidth]{Figures/volume_ca_104_scaleu2scaled5.eps}
\includegraphics[width=0.5\textwidth]{Figures/aver_moving_window6scaleu2scaled5.eps}\\
\includegraphics[width=0.5\textwidth]{Figures/flux_moving_window6scaleu2scaled5.eps}\\
\caption{The periodic (top left), time unit cells averaging (top right), and spatial location
(bottom), Eq. \ref{inlet:outlet:spatial:location}, calculated  volumetric mass transfer
coefficients. One can see that they all coincide. However, the periodic boundary conditions based
calculations produce the overestimated volumetric mass transfer
coefficient. \label{fig:periodic:ca1040}}
\end{figure}

@@

domain concentration through any formulation as they produce consistent results. However, the
result is the volumetric mass transfer coefficient will be overestimated using this
approach. Formulation of

@@

have indicator of how well the liquid slug is mixed. As well, for velocity patterns for $Ca\geq
0.7$ simulations with a few unit cells allow to calculate the volumetric mass transfer
coefficient based on the spatial location, not time averaged values used in all other
approaches. Finally, a sample of results was

### 2012-02-06 Text/paper.tex
Started a new document about definition of the mass transfer.
Need to keep the same Peclet number either with periodic boundary conditions (easiest case). Calculate critical velocities.
Started doing showcases for the second part of the paper.

\newcommand{\beq}{\begin{equation}}
\newcommand{\feq}{\end{equation}}
\newcommand{\beqstar}{\begin{equation*}}
\newcommand{\feqstar}{\end{equation*}}                                     
\newcommand{\beqal}{\begin{equation}\begin{aligned}}
\newcommand{\feqal}{\end{aligned}\end{equation}}
\newcommand{\beqalstar}{\begin{equation*}\begin{aligned}}
\newcommand{\feqalstar}{\end{aligned}\end{equation*}}
\newcommand{\vol}{k_L\,a}

@@

\subsection{Comparison between different boundary conditions}
The mass transfer coefficient by the definition of used in the paper of \citet{vanbaten} is as
follows:

>>>>

\newcommand{\beq}{\begin{equation}}
\newcommand{\feq}{\end{equation}}
\newcommand{\beqstar}{\begin{equation*}}
\newcommand{\feqstar}{\end{equation*}}                                     
\newcommand{\beqal}{\begin{equation}\begin{aligned}}
\newcommand{\feqal}{\end{aligned}\end{equation}}
\newcommand{\beqalstar}{\begin{equation*}\begin{aligned}}
\newcommand{\feqalstar}{\end{aligned}\end{equation*}}
\newcommand{\vol}{k_L\,a}
\newcommand{\lbubble}{L_{\mathrm{bubble}}}
\newcommand{\lunit}{L_{\mathrm{unit}}}
\newcommand{\lslug}{L_{\mathrm{slug}}}
\newcommand{\ububble}{U_{\mathrm{bubble}}}
\newcommand{\uliq}{U_{\mathrm{liq}}}
\newcommand{\ugas}{U_{\mathrm{gas}}}
\newcommand{\uoutlet}{U_{\mathrm{outlet}}}
\newcommand{\cbubble}{C_{\mathrm{bubble}}}
\newcommand{\cinlet}{C_{\mathrm{inlet}}}
\newcommand{\coutlet}{C_{\mathrm{outlet}}}
\newcommand{\coverall}{C_{\mathrm{overall}}}
\newcommand{\cstar}{C^{*}}
\newcommand{\cmedium}{C_{\mathrm{medium}}}
\newcommand{\volnondim}{\vol \frac{\lunit}{\ububble+\ugas}}
\newcommand{\omegaplus}{\omega_{+}}
\newcommand{\omegaminus}{\omega_{-}}
\newcommand{\holdup}{\varepsilon_{\mathrm{gas}}}

@@

\subsection{Comparison between different boundary conditions}
The mass transfer coefficient by the definition of used in the paper of \citet{vanbaten-circular} is
as
follows:
\begin{equation}
dd
\end{equation}
